---
title: "EDS241: Assignment 4"
author: "Charles Hendrickson"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document:
    toc: false
    number_sections: yes
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
--- 
  
``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}

# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3, 
                      echo = TRUE, message = FALSE, warning = FALSE)


# load packages
packages=c("stargazer", "here", "tidyr", "dplyr","stringr", "janitor", 
           "cowplot", "ggplot2", "tinytex", "datasets", "tibble", "openxlsx", "estimatr", "car", "readxl", "estimatr", "car", "AER")

# Load estimatr package
library(estimatr)

for (i in packages) {
  if (require(i,character.only=TRUE)==FALSE) {
    install.packages(i,repos='http://cran.us.r-project.org')
  }
  else {
    require(i,character.only=TRUE)
  }
}

#devtools::install_github('rstudio/rmarkdown')
options(scipen=10) # not scientific notation

```


# Data
```{r}
# Load the data
EU_sardines_data <- data.frame(read.csv("EU_sardines.csv"))
```


```{r}
# Take the log of volume_sold and price_euro_kg and create new columns for them
EU_sardines_data <- EU_sardines_data %>% 
  mutate(log_volume_sold_kg = log(volume_sold_kg)) %>% 
  mutate(log_price_euro_kg = log(price_euro_kg))

```


# Question (a)

(a) Estimate a bivariate regression of log(volume_sold_kg) on log(price euro_kg). What is the
price elasticity of demand for sardines? Test the null hypothesis that the price elasticity is equal to -1.

**The price elasticity of demand for sardines is -1.545**

**We reject the null hypothesis that the price elasticity is equal to -1 because our linear hypothesis test tells us that the p-value is less than 0.001 and is statistically significant.**

```{r}
# Bivariate regression of log(volume_sold_kg) on log(price euro_kg)
model_1 <- lm_robust(log_volume_sold_kg ~ log_price_euro_kg, data = EU_sardines_data)

# Get summary of the model
summary(model_1)

# Linear hypothesis test 
linearHypothesis(model_1, c("log_price_euro_kg = -1"), white.adjust = "hc2")

```

# Question (b)

(b) Like in Lecture 8 (see the IV.R script), we will use wind_m_s as an instrument for log(price_euro_kg). To begin, estimate the first-stage regression relating log(price_euro_kg) to wind_m_s. Interpret the estimated coefficient on wind speed. Does it have the expected sign? Also test for the relevance of the instrument and whether it is a “weak” instrument by reporting the proper F-statistic.

**The estimated coefficient on wind speed is 0.06735. This means that for every additional meter per second increase in wind speed, there is a 0.06735 percent increase in the log price of sardines. It does have the expected sign (positive) because increased wind speeds would make fishing effort more difficult, thus there should be a positive increase in the log price of sardines.**

**The F-statistic is 144.65, which means the wind_m_s instrument is relevant and non-weak (Lecture 9, slide 14).** 

```{r}
# Regress log(price_euro_kg) on wind_m_s
model_2 <- lm_robust(log_price_euro_kg ~ wind_m_s, data = EU_sardines_data)

# Get summary table of the regression coefficients 
summary(model_2)

# F-test for non-weak and relevant instruments 
linearHypothesis(model_2, c("wind_m_s = 0"), white.adjust = "hc2")
```

# Question (c)

(c) Estimate the TSLS estimator of the price elasticity of demand for sardines using wind_m_s
as an instrument for log(price_euro_kg). What is the estimated price elasticity of demand for
sardines?

**The two stage least squares (TSLS) regression of the price elasticity of demand for sardines using wind_m_s as an instrument for log(price_euro_kg) shows us that -1.08802 is the estimated price elasticity of demand for sardines.**

```{r}
# Two stage least squares regression 
tsls1 <- ivreg(log_volume_sold_kg ~ log_price_euro_kg | wind_m_s, data = EU_sardines_data)

# Summary of TSLS
summary(tsls1)

```

# Question (d)

(d) Repeat the exercise in (c), but include fixed effects for each year, month, and country. [Hint:
you can use the command “as.factor(country) + as.factor(year) +as.factor(month)” to the ivreg
function in R]. Report the estimated price elasticity of demand and the F-statistic testing for relevant and non-weak instruments.

**The estimated price elasticity of demand for sardines with fixed effects for each year, month, and country is -1.25004**

```{r}
# Completes two stage least regression with fixed effects for each year, month, and country
tsls2 <- ivreg(log_volume_sold_kg ~ log_price_euro_kg + 
               as.factor(country) + as.factor(year) + as.factor(month) | 
               wind_m_s + as.factor(country) + as.factor(year) + as.factor(month), 
               data = EU_sardines_data)

# Get summary table of TSLS
summary(tsls2)

```




**The F-statistic is 77.658, which means the instruments are relevant and non-weak (Lecture 9, slide 14).**



```{r}
# Can't get the f-statistic from ivreg
model_3 <- lm_robust(log_price_euro_kg ~ wind_m_s + as.factor(country) + as.factor(year) + as.factor(month), data = EU_sardines_data)

# F-test for non-weak and relevant instruments (Lecture 9, slides 13-14)
linear_hypoth_model_3 <- linearHypothesis(model_3, c("wind_m_s = 0"), white.adjust = "hc2")

linear_hypoth_model_3
```





































